Substructure Boosts to Dark Matter Annihilation from Sommerfeld Enhancement

### 2009-02-26 tex/somm.tex
Various Weiner sggestions incorporated

\begin{abstract}
I discuss the implications of the recently introduced Sommerfeld
enhancement of the dark matter annihilation cross section for the
detection of subhalos in the Galactic halo.  In addition to the boost
to the dark matter annihilation cross section from the high densities
of these subhalos with respect to the main halo, an additional boost
due to Sommerfeld enhancement results from the fact that they are
kinematically colder than the Galactic halo. If we further believe the
generic prediction of \CDM\ that in each subhalo there is an abundance
of substructure which is approximately self-similar to that of the
Galactic halo, then additional boosts coming from the density
enhancements of these small substructures and their small velocity
dispersions enhance the dark matter annihilation cross section even
further. I show that very large boost factors ($10^5$ -- $10^9$) are
obtained in a large class of models. The implications of these boost
factors for the detection of dark matter annihilation from dwarf
Speroidal galaxies in the Galactic halo are such that many of the
models proposed to explain the positron excess seen by the PAMELA
experiment are already marginally excluded.
\end{abstract}

@@

In addition to this larger boost due to the small
velocity dispersion in the subhalos, these subhalos themselves are
filled with an abundance of substructure which is approximately
self-similar to that of the main halo. Each of these even smaller
substructure has a larger density than the subhalo they are orbiting
in (quantified by the well established concentration-mass
relationships for \CDM\ halos
\cite{2001MNRAS.321..559B,2001ApJ...554..114E}) as well as a smaller
velocity dispersion. This opens up the possibility that the dark
matter annihilation signal in these already cold subhalos can be
boosted even further by contribution of their substructure. In this
paper we quantify the magnitude of this new substructure boost and
discuss its implications for the detection of dark matter annihilation
from subhalos in the Galactic halos.

@@

\subsection{Some details of the particle physics that do not particularly concern us here}\label{sec:partphys}

For the annihilation cross section without taking account of any
Sommerfeld enhancement we will adopt the fiducial value $\sigmaannv =
3 \times 10^{-26} $ cm$^3$ s$^{-1}$. The energy spectrum of the \DM\
annihilation depends on the model used to describe the \DM\ and its
interactions and it is outside of the scope of this paper to provide a
detailed treatment of this for different \DM\ models. Therefore, we
use an annihilation spectrum appropriate for neutralino dark matter,
which provides a reasonable fit to a variety of different annihilation
products. Since the energy spectrum contributes a multiplicative
factor to the \DM\ annihilation flux, the effect of different
annihilation spectra can be incorporated by simply multiplying our
results with the appropriate factor. The neutralino dark matter
annihilation spectrum we use has the form
\cite{1998APh.....9..137B,2004PhRvD..70j3529F}
\begin{equation}\label{eq:annspectrum}
\frac{\dd \Ngamma}{\dd E} = \frac{a_1}{\mdm} \left(\frac{E}{\mdm}\right)^{-3/2}\, \exp\left[-a_2\frac{E}{\mdm}\right]\, ,
\end{equation}
with parameters $a_1\sim 1$ and $a_2 \sim 10$, depending on the
details of the annihilation process
\cite{2004PhRvD..70j3529F}. Integrating this spectrum using a
threshold energy appropriate for \Fermi\, $\Eth = 4$ GeV, and choosing
a \DM\ mass, $\mdm = 700$ GeV, we find that
\begin{equation}
\int_{\Eth}^{\mdm}\dd E\, \sum_i \frac{\dd
\Ngammai}{\dd E} \approx 17\, ,
\end{equation}
with any differences due to variations in $(a_1,a_2)$ and $\mdm$
between different models leading to changes of a factor $\order(10)$.

@@

We can now write
\begin{equation}
\begin{split}
\lum(M) & = \somm\,\lumsmoothc(M) + \lumsub(M)\\
&= \left[ \somm + \frac{\lumsub(M)}{\lumsmoothc(M)}\right] \lumsmoothc(M)\, .
\end{split}
\end{equation}
We will from now on write the substructure boost as \boost(M). Then we
have the following
\begin{equation}\label{eq:boost}
\begin{split}
\boost(M) & \equiv \frac{\lumsub(M)}{\lumsmoothc(M)} = \frac{1}{\lumsmoothc(M)} \int \dd m \, \frac{\dd N}{\dd m} \lum(m)\\
&= \frac{1}{\lumsmoothc} \int \dd m\, \frac{\dd N}{\dd m} \left[ \somm(m) + \boost(m) \right] \lumsmoothc(m)\, .
\end{split}
\end{equation}
Here we have introduced the subhalo mass function $\dd N/\dd m$, which
gives the abundance of subhalos of mass $m$.

@@

We define the total boost for a given subhalo mass as
\begin{equation}\label{eq:totalboost}
\totalboost(M) = \somm(M) + \boost(M)\, ,
\end{equation}
in which $\somm(M)$ is the Sommerfeld enhancement for a halo of mass
$M$ and $\boost(M)$ is the substructure boost for a halo of that mass,
as defined in Eq.~(\ref{eq:boost}). This total boost for subhalos of
mass $M$ = $10^8$ \Msun\ and $M$ = $10^7$ \Msun\ calculated by
following the procedure outlined in the previous subsection is shown
in Figure \ref{fig:subboost} (\emph{top panels}). It is immediately
obvious that the total boost factors can get very large in certain
regions of the parameter space spanned by \mv/\mdm and $\alpha$:
boosts vary between $10^2$ to $10^9$. These masses correspond to the
typical masses of the classical and newly-discovered dwarf Spheroidals
(\dSphs) and as such their velocity dispersions are of the order
$10^{-5}$. The pattern of resonance regions and non-resonance regions
that appears for the pure Sommerfeld enhancement (as shown in Figure
\ref{fig:sommavg}) is also apparent in these plots of the total
boosts, however, it is noteworthy that the resonances seems to be more
highly peaked than they were in the pure Sommerfeld enhancement case,
and that the boost the lower half of the parameter space, \ie, for
smaller values of \mv/\mdm, are higher than one would expect from the
pure Sommerfeld enhancement combined with a boost factor from pure
density enhancements.

@@

\begin{figure*}[t]
\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 290 433 500
\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_sextans_65.eps}\\%%BoundingBox: 190 290 433 500
\includegraphics[clip=]{flux_umaii_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 290 433 500
\caption{Predicted \DM\ annihilation fluxes for various dwarf
Spheroidal satellite galaxies. The properties of the \dSphs\ used are
given in Table \ref{tab:dsphs}.}%
\label{fig:dsph}%
\end{figure*}

@@

Besides detection of \DM\ annihilation from the Galactic Center or the
diffuse component from the Galactic halo, there idea exists that the
best candidates for \DM\ annihilation detection might be the dwarf
Spheroidal satellite galaxies (\dSphs) of our Galaxy, since the dwarf
galaxies are believed to be the most dark matter dominated structures
in the Universe, and the \DM\ annihilation signal coming from them
would be much easier to distinguish from background sources than, \eg,
would be the case for \DM\ annihilation photons coming from the
Galactic Center. That \dSphs\ make good candidates for \DM\
annihilation detection has been much discussed before, especially for
the ``classical'', or pre-\SDSS, \dSphs\
\cite{2000PhRvD..61b3514B,2002PhRvD..66b3509T,2004PhRvD..70b3512B,2004PhRvD..69l3501E,2004PhRvD..70d3503P,2004PhRvL..93p1302H}. The
annihilation signal coming from Draco has been the subject of much
research in recent years because of its relative proximity to the
Earth
\cite{2002PhRvD..66b3509T,2004PhRvD..70d3503P,2006PhRvD..73f3510B,2006JCAP...03..003P,2007PhRvD..75b3513C,2007PhRvD..76l3509S,2007PhRvD..75h3526S,2008arXiv0812.1494P}. The
prospects for \DM\ annihilation detection from \dSphs\ has in general
been found to be rather poor, with detection with \Fermi\ being
unlikely \cite{2008arXiv0812.1494P}. Some of the \dSphs\ have already
been the subject of observational searches for \DM\ annihilation and
some limits on the \DM\ annihilation signal have already been set
\cite{2008ApJ...678..594W,2008ApJ...679..428A,2008APh....29...55A,2008PhRvD..78h7101D,2009ApJ...691..175A,2008arXiv0810.3561M}.

@@

It is clear from Figure \ref{fig:dsph} that the \DM\ annihilation flux
predicted to be coming from the \dSphs\ is very large in a large class
of models. For example, the flux predicted for Draco exceeds the upper
limit set on the \DM\ annihilation flux set by
Ref.~\cite{2008ApJ...679..428A} for all models parametrized by
\mv/\mdm\ and $\alpha$ by many orders of magnitude. We have to stress
here that our predicted fluxes are all for the particle physics
specific parameters given in Section \ref{sec:partphys}, \ie,
\sigmaannv\ = $3 \times 10^{-26}$ cm$^3$ s$^{-1}$, \mdm\ = 700 GeV and
a particular form for the \DM\ annihilation spectrum. To predict the
gamma-ray flux for a particular model with different values of these
particle physics parameters one is required to calculate the particle
physics parameter specific factor to the \DM\ annihilation flux and
appropriately rescale the values given here. Nevertheless, any model
that employs the Sommerfeld enhancement mechanism to predict the very
large annihilation cross sections necessary to account for the
PAMELA/ATIC excesses, and we are thinking here specifically about
models that try to fit the PAMELA data with very large boost factors
with respect to the fiducial annihilation cross section, give rise to
these very large predicted gamma-ray fluxes coming from the
\dSphs. Therefore, unless the boosts from substructure are somehow
suppressed or the assumptions about the subhalo mass function,
concentration-mass relation for subhalos, and/or any of the other
asumptions about subhalos beyond the resolution limits of the
state-of-the-art numerical simulations are violated for physical
subhalos, these models are in tension with the experimental limits on
the gamma-ray flux from \dSphs.

%dSphs good candidates for \DM\ annihilation:
%2000PhRvD..61b3514B,%Baltz/SUSY 2002PhRvD..66b3509T,%Tyler,Draco
%2004PhRvD..70b3512B,%Baltz,general 
%2004PhRvD..69l3501E,%Evans,general
%2004PhRvD..70d3503P,%Peirani, Draco, Sagit, M31, M87.
%2004PhRvL..93p1302H,%Hooper: Sagit: 511 keV
%2006PhRvD..73f3510B,%Bergstrom,Draco 
%2006JCAP...03..003P,%Profumo,Draco, SUSY 
%2007PhRvD..75b3513C,%Colafrancesco,Draco
%2007PhRvD..76l3509S,%Sanchez-Conde,Draco
%2007PhRvD..75h3526S,%StrigariI 
%2008ApJ...678..614S,%StrigariII
%2008arXiv0809.2781G,%Geha 
%2008arXiv0812.1494P,%Pieri, no Draco, Umi,Carina or Sextans with Fermi

%Limits:
%2008ApJ...678..594W,%Wood: Whipple: Draco and Umi
%2008ApJ...679..428A,%Albert et al, MAGIC: <10^-11
%2008APh....29...55A,%HESS: Sagg: 
%2008PhRvD..78h7101D,%Driscoll: Draco with STACEE: limit on the spectrum
%2009ApJ...691..175A,%Hess: Canis Major
%2008arXiv0810.3561M,%MAGIC: Willman 1 < 10^-12

%dSph modeling:
%2007PhRvD..75h3526S,%StrigariI
%2007ApJ...667L..53W,%Walker: vdisp profiles
%2008ApJ...672..904P,%Penarubia+CDM
%2009AJ....137.3109W,%Walker, EM contamination

@@

\begin{figure}
\centering
\includegraphics{hist_-2_-3_10_3.eps}
\caption{Total gamma-ray flux from subhalos in \VL\ II for 1,000 random
observers at 8.5 kpc from the center of the main \VL\ halo for
\mv/\mdm\ = 10$^{-2}$ and $\alpha = 10^{-3}$.}%\
\label{fig:VL23_3}%
\end{figure}

@@

We only cover some typical cases for the Sommerfeld enhancement here
which suffices to show the general trend. We consider the particle
physics parameters \mv/\mdm\ and $\alpha$ which give rise to the
smallest Sommerfeld enahcements in Figures \ref{fig:VL23_2} and
\ref{fig:VL23_3}, \ie, \mv/\mdm\ = 10$^{-2}$ and $\alpha$ = 0.001. The
left panel of Figure \ref{fig:VL23_2} shows the distribution of
gamma-ray fluxes from \DM\ annihilation for the subhalos, while the
right panel shows the distribution of the total boost factors. As can
be seen, the total boost factors are quite small: the typical boost
factor is only about 30 with the highest boost factors being about
100. The total gamma-ray flux coming from subhalos is then also quite
small in this case, as is shown in Figure \ref{fig:VL23_3}, in which
the total gamma-ray flux is shown for 1,000 random observer at a
distance of 8.5 kpc from the galactic center. At most observation
locations this total flux is about 10$^{-7}$ cm$^2$ s$^{-1}$, with the
high-end tail not extending further than about half an order of
magnitude from this. This is still a gamma-ray flux which is quite
large, showing that the cumulative signal gives rise to the largest
gamma-ray fluxes.

Figures \ref{fig:VL51_1} and \ref{fig:VL51_2} show the same for a
different extremum of the Sommerfeld enhancement: the case in which
\mv/\mdm\ = 10$^{-5}$ and $\alpha$ = 0.1. As we saw before, in this
case the Sommerfeld enhancement can reach very high values, and thus
we see that the typical \DM\ annihilation signal coming from \DM\
subhalos is large: the typical flux for one subhalo is equal to the
total flux in the scenario described in the previous paragraph. The
total boost factors for most of the halos are very large, of the order
of 10$^7$, which gives rise to this very large photon-flux. The total
predicted flux in this case is many orders of magnitude larger than
the total flux in the other extremum of the Sommerfeld
enhancement. Thus, the exact value of the parameters \mv/\mdm\ and
$\alpha$, upon which the Sommerfeld enhancement depends, matters a
great deal for the predicted gamma-ray flux.

@@

The recently proposed Sommerfeld enhancement to the \DM\ annihilation
cross section forces us to reconsider statements made in the past
about the expected signal of \DM\ anihilation from cold substructures
in the halo of the Galaxy. Since the Sommerfeld enhancement depends
criticically on the velocity dispersion of the substructure in
question, as well as on the details of the particle physics model used
to describe the attractive (or, possibly, repulsive) force that gives
rise to the Sommerfeld enhancement, the contributions from the
specific particle physics model and the details of the \DM\
distribution in the substructure can no longer be as cleanly separated
as they have been before. This means that in a scenario in which there
is a significant Sommerfeld enhancement, previous claims of ``particle
physics independent'' conclusions about the \DM\ annihilation signal
from substructure (\eg,
\cite{2007PhRvD..75h3526S,2008Natur.456...73S}) need to be
reconsidered in this new framework. This complicates the analysis but
it opens up new possibilities for detecting \DM\ annihilation and
learning about the properties of the dark sector. Given a detailed
knowledge of the density structure and velocity dispersion profile of
the dwarf Spheroidal galaxies, much can be learned if we are able to
detect the \DM\ annihilation signal coming from these \dSphs, by
considering the relative signal and using the results of
Sec.~\ref{sec:dsphs}, or by using the detailed velocity dispersion
profile of the \dSphs\ \cite{2009arXiv0902.0362R}.

>>>>

\begin{abstract}
I discuss the implications of the recently introduced Sommerfeld
enhancement of the dark matter annihilation cross section for the
detection of subhalos in the Galactic halo.  In addition to the boost
to the dark matter annihilation cross section from the high densities
of these subhalos with respect to the main halo, an additional boost
due to Sommerfeld enhancement results from the fact that they are
kinematically colder than the Galactic halo. If we further believe the
generic prediction of \CDM\ that in each subhalo there is an abundance
of substructure which is approximately self-similar to that of the
Galactic halo, then additional boosts coming from the density
enhancements of these small substructures and their small velocity
dispersions enhance the dark matter annihilation cross section even
further. I show that very large boost factors ($10^5$ -- $10^9$) are
obtained in a large class of models. The implications of these boost
factors for the detection of dark matter annihilation from dwarf
Speroidal galaxies in the Galactic halo are such that they outshine
the extragalactic background gamma-ray flux and should be detectable
by \Fermi.
\end{abstract}

@@

In addition to this larger boost due to the small
velocity dispersion in the subhalos, these subhalos themselves are
filled with an abundance of substructure which is approximately
self-similar to that of the main halo. Each of these even smaller
substructure has a larger density than the subhalo they are orbiting
in (quantified by the well established concentration-mass
relationships for \CDM\ halos
\cite{2001MNRAS.321..559B,2001ApJ...554..114E}) as well as a smaller
velocity dispersion. This opens up the possibility that the dark
matter annihilation signal in these already cold subhalos can be
boosted even further by contribution of their substructure.

In this paper we quantify the magnitude of this new substructure boost
and discuss its implications for the detection of dark matter
annihilation from subhalos in the Galactic halos. As we will show, the
total combined boost from Sommerfeld enhancement and substructure can
reach very large values (10$^5$ -- 10$^9$), and these boosts are only
weakly dependent on the assumptions that we have to make about the
abundance of substructure and the \DM\ density profile. We use these
total boosts together with published \DM\ density profiles of some of
the dwarf Spheroidal satellite galaxies of the Milky Way to predict
the total flux of gamma-rays from \DM\ annihilation that would be
detectable by Fermi and find that in most of the models that posit the
Sommerfeld enhancement the \DM\ annihilation signal from the dwarf
Spheroidals shines above the diffuse extragalactic gamma-ray
background, which is the only background signal at the high Galactic
lattitudes of these satellite galaxies. Finally, we also consider the
prospects for detecting smaller subhalos of the Galaxy by computing
annihilation fluxes of subhalos in the \VL\-2 simulation. We will see
that whether these subhalos will be detectable in the near future is
highly dependent on the exact details of the Sommerfeld enhancement.

@@

\subsection{Some details of the particle physics}\label{sec:partphys}

For the annihilation cross section without taking account of any
Sommerfeld enhancement we will adopt the fiducial value $\sigmaannv =
3 \times 10^{-26} $ cm$^3$ s$^{-1}$, and we choose a fiducial mass for
the \DM\ particle of \mdm\ = 700 GeV. The energy spectrum of the \DM\
annihilation depends on the model used to describe the \DM\ and its
interactions and it is outside of the scope of this paper to provide a
detailed treatment of this for different \DM\ models. Therefore, we
will adopt an annihilation spectrum for \DM\ annihilation into photons
appropriate for the newer models that were proposed to explain the
PAMELA/ATIC results. Specifically, we use a gamma ray spectrum from a
cascade annihilation model. In these models the \DM\ annihilates into
a new light degree of freedom which then in turn decays into leptons
and photons. The injection spectrum per \DM\ annihilation is given by
\cite{2009arXiv0901.2926M}
\begin{equation}\label{eq:annspectrum}
\begin{split}
\frac{\dd \Ngamma}{\dd E} = &\frac{\alpha_{\mbox{{\footnotesize EM}}}}{\pi} \frac{1+(1-E/\mdm)^2}{E}\\
&\times\left\{-1 + \ln\left(4(1-E/\mdm)\right) -2\ln\left(\frac{m_l}{\mdm}\right)\right\}\, ,
\end{split}
\end{equation}
in which $\alpha_{\mbox{{\footnotesize EM}}}$ is the electromagnetic
coupling constant and $m_l$ is the mass of the leptons created in the
process $\chi\chi \rightarrow l^+l^-\gamma$. Integrating this spectrum
using an energy range appropriate for \Fermi\ we find that
\begin{equation} 
\int_{4 \mbox{{\footnotesize GeV}}}^{250 \mbox{{\footnotesize GeV}}}\dd E\, \frac{\dd
\Ngamma}{\dd E} \approx 0.5\, .
\end{equation}
We find a similar number if we integrate the spectrum due to one
cascade step. It is interesting to note that these photon spectra have
fewer photons in them, about an order of magnitude less, than the
annihilation spectra due to neutralino annihilation which are often
used \cite{1998APh.....9..137B,2004PhRvD..70j3529F}. Of course, since
the energy spectrum contributes a multiplicative factor to the \DM\
annihilation flux, the effect of different annihilation spectra can be
incorporated by simply multiplying our results with the appropriate
factor.

@@

We can now write
\begin{equation}\label{eq:totboost}
\begin{split}
\lum(M) & = \somm\,\lumsmoothc(M) + \lumsub(M)\\
&= \left[ \somm + \frac{\lumsub(M)}{\lumsmoothc(M)}\right] \lumsmoothc(M)\, .
\end{split}
\end{equation}
We will from now on write the substructure boost as \boost(M). Then we
have the following
\begin{equation}\label{eq:boost}
\begin{split}
\boost(M) & \equiv \frac{\lumsub(M)}{\lumsmoothc(M)} = \frac{1}{\lumsmoothc(M)} \int \dd m \, \frac{\dd N}{\dd m} \lum(m)\\
&= \frac{1}{\lumsmoothc} \int \dd m\, \frac{\dd N}{\dd m} \left[ \somm(m) + \boost(m) \right] \lumsmoothc(m)\, .
\end{split}
\end{equation}
Here we have introduced the subhalo mass function $\dd N/\dd m$, which
gives the abundance of subhalos of mass $m$.

@@

We define the total boost for a given subhalo mass as (see \eqnname\
[\ref{eq:totboost}])
\begin{equation}\label{eq:totalboost}
\totalboost(M) = \somm(M) + \boost(M)\, ,
\end{equation}
in which $\somm(M)$ is the Sommerfeld enhancement for a halo of mass
$M$ and $\boost(M)$ is the substructure boost for a halo of that mass,
as defined in Eq.~(\ref{eq:boost}). This total boost for subhalos of
mass $M$ = $10^8$ \Msun\ and $M$ = $10^7$ \Msun\ calculated by
following the procedure outlined in the previous subsection is shown
in Figure \ref{fig:subboost} (\emph{top panels}). It is immediately
obvious that the total boost factors can get very large in certain
regions of the parameter space spanned by \mv/\mdm and $\alpha$:
boosts vary between $10^2$ to $10^9$. These masses correspond to the
typical masses of the classical and newly-discovered dwarf Spheroidals
(\dSphs) and as such their velocity dispersions are of the order
$10^{-5}$. The pattern of resonance regions and non-resonance regions
that appears for the pure Sommerfeld enhancement (as shown in Figure
\ref{fig:sommavg}) is also apparent in these plots of the total
boosts, however, it is noteworthy that the resonances seems to be more
highly peaked than they were in the pure Sommerfeld enhancement case,
and that the boost the lower half of the parameter space, \ie, for
smaller values of \mv/\mdm, are higher than one would expect from the
pure Sommerfeld enhancement combined with a boost factor from pure
density enhancements.

@@

\begin{figure*}[t]
\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 290 433 500
\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_sextans_65.eps}\\%%BoundingBox: 190 290 433 500
\includegraphics[clip=]{flux_umaii_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 290 433 500
\caption{Predicted \DM\ annihilation fluxes for various dwarf
Spheroidal satellite galaxies. The background flux of extragalactic
photons in the relevant energy range is given under the name of the
dwarf Spheroidal. The properties of the \dSphs\ used are given in
Table \ref{tab:dsphs}.}%
\label{fig:dsph}%
\end{figure*}

@@

Besides detection of \DM\ annihilation from the Galactic Center or the
diffuse component from the Galactic halo, there idea exists that the
best candidates for \DM\ annihilation detection might be the dwarf
Spheroidal satellite galaxies (\dSphs) of our Galaxy, since the dwarf
galaxies are believed to be the most dark matter dominated structures
in the Universe, and the \DM\ annihilation signal coming from them
would be much easier to distinguish from background sources than, \eg,
would be the case for \DM\ annihilation photons coming from the
Galactic Center. That \dSphs\ make good candidates for \DM\
annihilation detection has been much discussed before, especially for
the ``classical'', or pre-\SDSS, \dSphs\
\cite{2000PhRvD..61b3514B,2002PhRvD..66b3509T,2004PhRvD..70b3512B,2004PhRvD..69l3501E,2004PhRvD..70d3503P,2004PhRvL..93p1302H}. The
annihilation signal coming from Draco has been the subject of much
research in recent years because of its relative proximity to the
Earth
\cite{2002PhRvD..66b3509T,2004PhRvD..70d3503P,2006PhRvD..73f3510B,2006JCAP...03..003P,2007PhRvD..75b3513C,2007PhRvD..76l3509S,2007PhRvD..75h3526S,2008arXiv0812.1494P}. The
prospects for \DM\ annihilation detection from \dSphs\ has in general
been found to be rather poor, with detection with \Fermi\ being
unlikely \cite{2008arXiv0812.1494P,2009arXiv0902.4330P}. Some of the
\dSphs\ have already been the subject of observational searches for
\DM\ annihilation and some limits on the \DM\ annihilation signal have
already been set
\cite{2008ApJ...678..594W,2008ApJ...679..428A,2008APh....29...55A,2008PhRvD..78h7101D,2009ApJ...691..175A,2008arXiv0810.3561M}.

@@

It is clear from Figure \ref{fig:dsph} that the \DM\ annihilation flux
predicted to be coming from the \dSphs\ is very large in a large class
of models. We have to stress here that our predicted fluxes are all
for the particle physics specific parameters given in Section
\ref{sec:partphys}, \ie, \sigmaannv\ = $3 \times 10^{-26}$ cm$^3$
s$^{-1}$, \mdm\ = 700 GeV and a particular form for the \DM\
annihilation spectrum. To predict the gamma-ray flux for a particular
model with different values of these particle physics parameters one
is required to calculate the particle physics parameter specific
factor to the \DM\ annihilation flux and appropriately rescale the
values given here. In order to compare these signals to the expected
background flux of extragalactic gamma rays (which is the only
relevant background since these \dSphs\ are at high Galactic
lattitude), we have integrated the extragalactic gamma ray flux as
measured by EGRET over the energy range between 4 GeV -- 250 GeV and
multiplied it with the solid angle that the \dSphs\ span on the
sky. The extragalactic gamma ray flux is given by a power law
\cite{1998ApJ...494..523S}
\begin{equation}
\frac{\dd \Ngamma}{\dd E \,\dd A \, \dd t\,\dd \Omega} = k\,\left(\frac{E}{E_0}\right)^{-\alpha}\, ,
\end{equation}
where $k = 7.3 \times 10^{-9}$, $\alpha$ = 2.10, and $E_0 = 0.45$
GeV. We have included the background flux of extragalactic photons in
Figure \ref{fig:dsph}. This background flux is in all cases of the
order of 10$^{-13}$ photons cm$^{-2}$ s$^{-1}$, such that the \DM\
annihilation boosted by substructure and Sommerfeld enhancement
outshine this background in all of the considered models, and by many
orders of magnitude in most models. Given that the flux above 100 MeV
over a 55 days time-span needed to separate a point source of \DM\
annihilation from the background by \Fermi\ to give a 5$\sigma$
detection is about 10$^{-8}$ photons cm$^{-1}$ s$^{-1}$ and that the
necessary fluxes scale roughly as $t_{\mbox{{\footnotesize
exp}}}^{-1/2}$ \cite{2008JCAP...07..013B}, the \DM\ annihilation of
these \dSphs\ should be detectable by \Fermi\ in many of the models
parameterized by ($\mv/\mdm,\alpha$). We conclude that the prospects
for detecting \DM\ annihilation from the \dSphs\ with \Fermi\ are very
good in models that include a significant Sommerfeld enhancement.

@@

\begin{figure}
\centering
\includegraphics{hist_-2_-3_10_3.eps}
\caption{Total gamma-ray flux from subhalos in \VL\ II for 1,000
random observers at 8.5 kpc from the center of the main \VL\ halo for
\mv/\mdm\ = 10$^{-2}$ and $\alpha = 10^{-3}$. For comparison we show
the total gamma-ray flux from subhalos in the absence of Sommerfeld
enhancement as the dashed histogram and the top horizontal axis.}%\
\label{fig:VL23_3}%
\end{figure}

@@

We only cover some typical cases for the Sommerfeld enhancement here
which suffices to show the general trend. We consider the particle
physics parameters \mv/\mdm\ and $\alpha$ which give rise to the
smallest Sommerfeld enahcements in Figures \ref{fig:VL23_2} and
\ref{fig:VL23_3}, \ie, \mv/\mdm\ = 10$^{-2}$ and $\alpha$ = 0.001. The
left panel of Figure \ref{fig:VL23_2} shows the distribution of
gamma-ray fluxes from \DM\ annihilation for the subhalos, while the
right panel shows the distribution of the total boost factors. As can
be seen, the total boost factors are quite small: the typical boost
factor is only about 30 with the highest boost factors being about
100. The total gamma-ray flux coming from subhalos is then also quite
small in this case, as is shown in Figure \ref{fig:VL23_3}, in which
the total gamma-ray flux is shown for 1,000 random observers at a
distance of 8.5 kpc from the galactic center. At most observation
locations this total flux is about 10$^{-7}$ cm$^2$ s$^{-1}$, with the
high-end tail not extending further than about half an order of
magnitude from this. For comparison the dashed histogram in Figure
\ref{fig:VL23_3} shows the same distribution of total gamma-ray flux
for different random observers in the absence of any Sommerfeld
enhancement. This confirms that the total gamma-ray flux with
Sommerfeld enhancement is about an order of magnitude larger than
without Sommerfeld enhancement.

Figures \ref{fig:VL51_1} and \ref{fig:VL51_2} show the same for a
different extremum of the Sommerfeld enhancement: the case in which
\mv/\mdm\ = 10$^{-5}$ and $\alpha$ = 0.1. As we saw before, in this
case the Sommerfeld enhancement can reach very high values, and thus
we see that the typical \DM\ annihilation signal coming from \DM\
subhalos is large: the typical flux for one subhalo is equal to the
total flux in the scenario described in the previous paragraph. The
total boost factors for most of the halos are very large, of the order
of 10$^7$, which gives rise to this very large photon-flux. The total
predicted flux in this case is many orders of magnitude larger than
the total flux in the other extremum of the Sommerfeld enhancement and
about seven orders of magnitude larger than the total gamma-ray flux
in the absence of any Sommerfeld enhancement. Thus, the exact value of
the parameters \mv/\mdm\ and $\alpha$, upon which the Sommerfeld
enhancement depends, matters a great deal for the predicted gamma-ray
flux.

@@

The recently proposed Sommerfeld enhancement to the \DM\ annihilation
cross section forces us to reconsider statements made in the past
about the expected signal of \DM\ anihilation from cold substructures
in the halo of the Galaxy. Since the Sommerfeld enhancement depends
criticically on the velocity dispersion of the substructure in
question, as well as on the details of the particle physics model used
to describe the attractive (or, possibly, repulsive) force that gives
rise to the Sommerfeld enhancement, the contributions from the
specific particle physics model and the details of the \DM\
distribution in the substructure can no longer be as cleanly separated
as they have been before. This means that in a scenario in which there
is a significant Sommerfeld enhancement, previous claims of ``particle
physics independent'' conclusions about the \DM\ annihilation signal
from substructure (\eg,
\cite{2007PhRvD..75h3526S,2008Natur.456...73S}) need to be
reconsidered in this new framework. This complicates the analysis but
it opens up new possibilities for detecting \DM\ annihilation and
learning about the properties of the dark sector. As we showed above,
the combination of the Sommerfeld enhancement and the substructure
that is believed to exists down to small scales leads to very large
boost factors because of the combined effect of the higher densities
of small subhalos as well as the fact that they are kinematically
colder than the larger halos. This leads to predictions for the \DM\
annihilation gamma-ray flux that are much larger than was previously
believed. As we showed in Sec.~\ref{sec:dsphs}, in models in which
there is a significant Sommerfeld enhancement at the level of the
Galactic halo and its subhalos, the \DM\ annihilation signal from the
dwarf Spheroidals outshines the extragalactic background flux and the
predicted gama-ray fluxes are so large that they should be detectable
by \Fermi. Given a detailed knowledge of the density structure and
velocity dispersion profile of the dwarf Spheroidal galaxies, much can
then be learned if we are able to detect the \DM\ annihilation signal
coming from these \dSphs, by considering the relative signal and using
the results of Sec.~\ref{sec:dsphs}, or by using the detailed velocity
dispersion profile of the \dSphs\ \cite{2009arXiv0902.0362R}.
